title: "Physical Activity" - Project 1
output: html_document
Created by Bonnie Mohnsen

Load and read data
```{r}
unzip("activity.zip")
dat <- read.csv("activity.csv", sep = ",", header = T)
```

Include libraries and set up
```{r}
library(ggplot2)
echo = TRUE
```	 
What is the total number of steps taken per day?
```{r}
OrigTotalSteps <- tapply(dat$steps, dat$date, FUN=sum, na.rm=TRUE)
OrigTotalSteps
```  
What is mean total number of steps taken per day?
```{r}
OrigMean <- mean(OrigTotalSteps, na.rm=TRUE)
OrigMean
```  
What is median total number of steps taken per day?
```{r}
OrigMedian <- median(OrigTotalSteps, na.rm=TRUE)
OrigMedian
```     
Original histogram
```{r}
hist(OrigTotalSteps, breaks = 30, col = "Lightblue", main = "Histogram of Daily Total Steps", xlab ="Total Steps/Day", ylab = "Frequency")
```     
What is the average daily activity pattern?
```{r}
OrigAverage <- aggregate(x=list(steps=dat$steps), by=list(interval=dat$interval), FUN=mean, na.rm=TRUE)
```  
Make a time series plot (i.e., type = "1") of the 5-min interval (x) and average step across all days (y-axis)
```{r}
ggplot(data=OrigAverage, aes(x=interval, y=steps)) + geom_line(color = "blue") + xlab("5-minute interval") + ylab("average steps") + ggtitle("Activity Pattern")
```     
Which 5-minute interval, on average across all days in the dataset, contains the maximum number of steps?
```{r}
OrigAverageMaxTime <- OrigAverage[which.max(OrigAverage$steps),]
OrigAverageMaxTime
```    
Calculate and report, the total number of rows with NAs
```{r}
NoMissingData <- is.na(dat$steps)
table(NoMissingData)
```    
Replace missing values using mean for that 5 minute interval to create Tidy Data Set
```{r}
Tidy <- dat
for (i in 1:nrow(Tidy)) {
     if (is.na(Tidy$steps[i])) {
          Tidy$steps[i] <- OrigAverage[which(Tidy$interval[i] == OrigAverage$interval), ]$steps
}}
```    
Tidy  - total number of steps per day
```{r}
TidyTotalSteps <- tapply(Tidy$steps, Tidy$date, FUN=sum)
TidyTotalSteps
```    
Tidy  -  histogram
```{r}
hist(TidyTotalSteps, breaks = 30, col = "Red", main = "Histogram of Tidy Daily Total Steps", xlab ="Total Steps/Day", ylab = "Frequency")
```    
Tidy mean and median
```{r}
TidyMean <- mean(TidyTotalSteps, na.rm=TRUE)
TidyMean
TidyMedian <- median(TidyTotalSteps, na.rm=TRUE)
TidyMedian
```  
Difference between Original mean and Tidy Mean
```{r}
OrigMean - TidyMean
```   
Difference between Original Median and Tidy Median
```{r}
OrigMedian - TidyMedian
```    
Tidy - differences between activity patterns on weekdays and weekends. Create a new factor variable(weekend (TRUE) or weekday)
```{r}
Tidy$day <- weekdays(as.Date(Tidy$date))
Tidy$Weekend <- with(Tidy, day == "Sunday") | with(Tidy, day == "Saturday")
```    
Tidy - Aggregate data by Weekdays (Weekend or Weekday) and Interval
```{r}
TidyAverage <- aggregate(x=list(steps=Tidy$steps), by=list(interval=Tidy$interval,Tidy$Weekend), FUN=mean)
```    
Tidy - Average steps: 0 - weekdays and 1 - weekends

```{r}
ggplot(TidyAverage, aes(interval, steps)) + geom_line() + facet_grid(Group.2 ~ .) + xlab("5-Minutes Intervals") + ylab("Average steps")
```    